Parallel Information Set Generation for Kriegspiel

### 2012-07-11 sbac/paper.tex
Changed the style according to the conf web site guidelines.. They must have changed the guidelines recently

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}

@@

\Section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{thielscher12hyperplay}.

@@

%\Section{Information Set Generation}
\Section{Background: Kriegspiel}
\label{Background}
Information Set Generation is a widely adopted operation used in game trees search. Here, we present our techniques and results in the context of a particular game - Kriegspiel. However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
%\Section{Background: Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

@@

\Section{Information Set Generation}
\label{info}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

\Section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

\Section{Performance Results}\label{Results}

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\vspace{-0.1in}
\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
\label{speedups}
\vspace{-0.3in}
\end{table}
In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance. Next, we show that our implementation performs equally well for instances from actual games.
\subsection{Analysis using Random Instance}
The performance results reported in
this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation

@@

\begin{figure*}[!ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.2in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.2in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.2in]{plots/dep12.pdf}
\label{dep12}
} 
\vspace{-0.1in}
\caption{Grain Size Control for different problem sizes}
%\vspace{-0.1in}
\label{fig:dep}
\vspace{-0.1in}
\end{figure*}

@@

\Section{Related Work}
\label{Related}
An overview of the theory of imperfect information games and other basic topics in game theory can be found
in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
found in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash
equilibrium can be found in time that is polynomial in the number of nodes in the game tree~\cite{koller94fast}.  In fact, they found
that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
computer systems tend to utilize the concept of information sets or belief states in some form or another.  A common
strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
Scrabble~\cite{sheppard02world}).

@@

\Section{Conclusion and Future Work}
We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
ordering heuristics.

@@

\bibliographystyle{latex8}
\bibliography{paper}

>>>>

%\documentclass[times, 10pt,twocolumn]{article}
%\documentclass[conference]{IEEEtran}
\documentclass[10pt, conference, compsocconf]{IEEEtran} 
%\usepackage{latex8}
%\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}

@@

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{thielscher12hyperplay}.

@@

%\Section{Information Set Generation}
\section{Background: Kriegspiel}
\label{Background}
Information Set Generation is a widely adopted operation used in game trees search. Here, we present our techniques and results in the context of a particular game - Kriegspiel. However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
%\Section{Background: Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

@@

\section{Information Set Generation}
\label{info}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

\section{Performance Results}\label{Results}

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\vspace{-0.1in}
\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
\label{speedups}
\vspace{-0.3in}
\end{table}
In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance. Next, we show that our implementation performs equally well for instances from actual games.
\subsection{Analysis using Random Instance}
The performance results reported in
this section were obtained by running our program on Blueprint with up to 1024 processors. This system was composed of 120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was decomissioned in June 2011 and is no longer in operation.

@@

\begin{figure*}[!ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\vspace{-0.1in}
\caption{Grain Size Control for different problem sizes}
%\vspace{-0.1in}
\label{fig:dep}
\vspace{-0.1in}
\end{figure*}

@@

\section{Related Work}
\label{Related}
An overview of the theory of imperfect information games and other basic topics in game theory can be found
in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
found in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash
equilibrium can be found in time that is polynomial in the number of nodes in the game tree~\cite{koller94fast}.  In fact, they found
that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
computer systems tend to utilize the concept of information sets or belief states in some form or another.  A common
strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
Scrabble~\cite{sheppard02world}).

@@

\section{Conclusion and Future Work}
We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
ordering heuristics.

@@

\bibliographystyle{IEEEtran}
\bibliography{paper}
